Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2015-08-18 3dpas_intro.tex
tweaks.

\author{Shantenu Jha\affil{1,2}\corrauth\, Neil Chue Hong\affil{2},
  Simon Dobson\affil{3}, Daniel S. Katz\affil{4,5},
  Andre~Luckow\affil{1}, Omer Rana\affil{6}, Yogesh Simmhan\affil{7}}

\address{\affilnum{1}Rutgers University\break
\affilnum{2}University of Edinburgh\break
\affilnum{3}University of St.~Andrews\break
\affilnum{4}University of Chicago\break
\affilnum{5}Argonne National Laboratory\break
\affilnum{6}Cardiff University\break
\affilnum{7}Indian Institute of Science\break
}

@@

The 3DPAS Research Theme (funded by the eSI, Edinburgh and US NSF)
investigated applications that have these characteristics; in other
words, it operated at the triple point of  {\it distributed} and {\it dynamic} and
{\it data-intensive} (D3) attributes.  This document represents the conclusions
of the 3DPAS Theme.
\end{abstract}
\maketitle

@@

\begin{table}
  \begin{scriptsize}
    \begin{center}
      \caption{Coordination approaches used by the applications\label{table:progSystems}}
      \begin{tabular}{|p{6.3cm}|p{7.2cm}|}
        \hline Application  & Current approaches \\
        \hline
        \hline NGS Analytics & pipelines that ship with
        sequencers~\cite{illumina}; workflows~\cite{Taverna} \\
        \hline ATLAS & custom algorithms; job submission tied to WLCG \\
        \hline LSST & custom layer; scripted workflows; queries \\
        \hline SOA Astronomy & web services~\cite{barker} \\
        \hline Cosmic Microwave Background & \textit{ad hoc} \\
        \hline Sensor Network Application & focused on data collection \\
        \hline Climate & \textit{ad hoc} \\
        \hline Interactive Exploration of Environmental Data &
        visualization tools~\cite{blower2} \\
        \hline Power Grids & stream processing~\cite{Simmhan:sciencecloud:2011}; data mining~\cite{Yin:mapreduce:2012} \\
        \hline Fusion (ITER) & scripted pipeline~\cite{adios} \\
        \hline Industrial Incident Notification and Response &
        agent-based~\cite{emergency4}  \\
        \hline MODIS Data Processing & scripted pipeline~\cite{modis:ipdps:2010} \\
%        \hline Floating Sensors & \textit{ad hoc} \\
        \hline Distributed Network Intrusion Detection & monolithic
        packages \\
        \hline
      \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}

@@

This section brings together some key observations
from the applications and infrastructure
analysis carried out in this paper.
%as take away messages. 
Section~\ref{sec:discussion}
provides a discussion of dynamic and distributed characteristics of
data associated with the applications we analyze in this work,
attempting to infer more general characteristics that %can be applied
may apply
to other similar applications.  We also suggest two metrics that may
suggest two metrics that may
be associated with such applications, primarily as a basis to compare
how compute and data processing (in terms of their number and
distribution) may be used to find common features between them. Such
feature analysis is useful, we believe, for two main reasons: (i) to
identify similarities between the applications covered in
\S\ref{sec:scenarios} with others that are in the design phase or
under development, so that common requirements can be identified; and
(ii) to identify gaps in current software tools and programming
systems that such applications use, thereby leading to the
development of new tools and systems. As our application and
infrastructure coverage is, by necessity, limited, identifying common
features also enables the lessons learned to be generalized and
applied more widely. We subsequently discuss a set of architectures
for D3 science, and pose some questions we believe are pertinent as
the community begins considering the next-generation
architectures for D3 applications. In
\S\ref{sec:conclusion} we conclude by providing a number of
observations to inform applications and
cyberinfrastructure developers, based on common trends found in our study.

@@

From \S\ref{sec:distTypes} and
\figurename~\ref{fig:figures_application}, we find that applications
with distributed data differ based on the amount or volume of data
distributed, the degree of distribution (scattered over one site or
many sites), and the dynamic aspects associated with distribution,
e.g., latency and relative scheduling flexibility (and thus tolerance
or constraints on data transfer/aggregation).  Applications processing
the largest volumes of data (O(PB)) typically run on specialized data
cyberinfrastructure (e.g., the Atlas LHC Grid). Applications such as
Fusion and NGS that run on smaller data sets mainly utilize HPC
environments. On the other end of the spectrum are highly dynamic
applications that process sensor data and make near-realtime
decisions based on streaming data, e.g., the power grid and sensor
network applications. These applications process much less data;
often, sensor data is also archived and aggregated for historical
analyses, which leads to higher data volumes. \katznote{not sure about the `other' part of this...}

@@

>>>>

\author{Shantenu Jha\affil{1}\corrauth\, Daniel S. Katz\affil{4,5},
  Andre~Luckow\affil{1}, Omer Rana\affil{6}, Yogesh Simmhan\affil{7}, Neil Chue
  Hong\affil{2}, Simon Dobson\affil{3}}

\address{\affilnum{1}Rutgers University\break
\affilnum{2}University of Edinburgh\break
\affilnum{3}University of St.~Andrews\break
\affilnum{4}University of Chicago\break
\affilnum{5}Argonne National Laboratory\break
\affilnum{6}Cardiff University\break
\affilnum{7}University of Southern California\break
}

@@

% The 3DPAS Research Theme (funded by the eSI, Edinburgh and US NSF)
% investigated applications that have these characteristics; in other
% words, it operated at the triple point of  {\it distributed} and {\it dynamic} and
% {\it data-intensive} (D3) attributes.  This document represents the conclusions
% of the 3DPAS Theme.

\end{abstract}
\maketitle

@@

\begin{table}
  \begin{scriptsize}
    \begin{center}
      \caption{Coordination approaches used by the applications\label{table:progSystems}}
      \begin{tabular}{|p{6.3cm}|p{7.2cm}|}
        \hline Application  & Current approaches \\
        \hline
        \hline NGS Analytics & pipelines that ship with
        sequencers~\cite{illumina}; workflows~\cite{Taverna} \\
        \hline ATLAS & custom algorithms; job submission tied to WLCG \\
        \hline LSST & custom layer; scripted workflows; queries \\
        \hline SOA Astronomy & web services \\
        \hline Cosmic Microwave Background & \textit{ad hoc} \\
        \hline Sensor Network Application & focused on data collection \\
        \hline Climate & \textit{ad hoc} \\
        \hline Interactive Exploration of Environmental Data &
        visualization tools \\
        \hline Power Grids & stream processing; data mining \\
        \hline Fusion (ITER) & scripted pipeline \\
        \hline Industrial Incident Notification and Response &
        agent-based  \\
        \hline MODIS Data Processing & scripted pipeline \\
%        \hline Floating Sensors & \textit{ad hoc} \\
        \hline Distributed Network Intrusion Detection & monolithic
        packages \\
        \hline
      \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}

@@

This section brings together some key observations
from the applications and infrastructure
analysis carried out in this paper.
%as take away messages. 
Section~\ref{sec:observations}
summarize the dynamic and distributed characteristics of
data associated with the applications we've studied in this work,
attempting to infer more general characteristics that %can be applied
may apply
to other similar applications.  In \S\ref{sec:discussion}, we
suggest two metrics that may
be associated with such applications, primarily as a basis to compare
how compute and data processing (in terms of their number and
distribution) may be used to find common features between them. Such
feature analysis is useful, we believe, for two main reasons: (i) to
identify similarities between the applications covered in
\S\ref{sec:scenarios} with others that are in the design phase or
under development, so that common requirements can be identified; and
(ii) to identify gaps in current software tools and programming
systems that such applications use, thereby leading to the
development of new tools and systems. As our application and
infrastructure coverage is, by necessity, limited, identifying common
features also enables the lessons learned to be generalized and
applied more widely. We subsequently discuss a set of architectures
for D3 science, and pose some questions we believe are pertinent as
the community begins considering the next-generation
architectures for D3 applications. In
\S\ref{sec:conclusion} we conclude by providing a number of
observations to inform applications and
cyberinfrastructure developers, based on common trends found in our study.

@@

From \S\ref{sec:distTypes} and
\figurename~\ref{fig:figures_application}, we find that applications
with distributed data differ based on the amount or volume of data
distributed, the degree of distribution (scattered over one site or
many sites), and the dynamic aspects associated with distribution,
e.g., latency and relative scheduling flexibility (and thus tolerance
or constraints on data transfer/aggregation).  Applications processing
the largest volumes of data (O(PB)) typically run on specialized data
cyberinfrastructure (e.g., the Atlas LHC Grid). Applications such as
Fusion and NGS that run on smaller data sets mainly utilize HPC
environments. On the other end of the spectrum are highly dynamic
applications that process sensor data and make near-realtime
decisions based on streaming data, e.g., the power grid and sensor
network applications. These applications process much less data;
often, sensor data is also archived and aggregated for historical
analyses, which leads to higher data volumes.

@@

Now that we have discussed dynamic data and distributed data in the
context of specific applications, we can use this experience to
understand the characteristics that can be associated with D3 science
applications, and the characteristics of a computational
infrastructure that may be most suited for D3 science
applications. 
We believe these are general
trends that can be applied to other applications we have not
considered.
